Supplementary Materials

### 2023-09-03 reports/state_figures.Rmd
update figs

lagged_fig_a <- subset %>%
    filter(state %in% states_with_lag & biweek <= 25 & biweek >=6) %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill = 'Probabilistic Bias Analysis'),
             alpha = .8,
             show.legend=FALSE) +
  geom_ribbon(aes(x = date,
                  fill = 'Covidestim',
                  ymin = infections.lo,
                  ymax = infections.hi),
              alpha = .6) +
  # geom_vline(xintercept=end_old_model,
  #            linetype=2,
  #            color="darkred",
  #            linewidth=.5) +
  facet_wrap(~name, ncol=3, scales = "free",
              labeller= labeller(
                vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(axis.text.x=element_text(size=7),
          axis.text.y = element_text(size=4.5),
          legend.position = "top",
          legend.text=element_text(size=12),
          strip.text.x=element_text(size=11,
                                    face='plain',
                                    color='white')) +
  theme(plot.title = element_text(face="bold",
                                 hjust = 0, 
                                 size = 13,
                                 margin=margin(0,0,0,0))) +
  scale_fill_manual(values =comp_state_pal, name='')  +
  labs(y = "Estimated Infections",
       title = paste0(
         "(a) Estimated Infections by State in States with a Lag")) +
  guides(color = guide_legend(override.aes = list(linewidth =2),
                              nrow=2,
                              byrow=TRUE))
lagged_fig_a

>>>>

lagged_fig_a <- subset %>%
    filter(state %in% states_with_lag & biweek <= 25 & biweek >=6) %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill = 'Probabilistic Bias Analysis'),
             alpha = .8,
             show.legend=FALSE) +
  geom_ribbon(aes(x = date,
                  fill = 'Covidestim',
                  ymin = infections.lo,
                  ymax = infections.hi),
              alpha = .6) +
  # geom_vline(xintercept=end_old_model,
  #            linetype=2,
  #            color="darkred",
  #            linewidth=.5) +
  facet_wrap(~name, ncol=3, scales = "free",
              labeller= labeller(
                vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(
    #date_breaks = "3 months", 
    breaks=ymd("2021-03-12", "2021-05-21", "2021-07-30", "2021-10-08"),
               date_labels = "%b %Y") +
  theme_c(axis.text.x=element_text(size=7),
          axis.text.y = element_text(size=4.5),
          legend.position = "top",
          legend.text=element_text(size=12),
          strip.text.x=element_text(size=11,
                                    face='plain',
                                    color='white')) +
  theme(plot.title = element_text(face="bold",
                                 hjust = 0, 
                                 size = 13,
                                 margin=margin(0,0,0,0))) +
  scale_fill_manual(values =comp_state_pal, name='')  +
  labs(y = "Estimated Infections",
       title = paste0(
         "(a) Estimated Infections by State in States with a Lag")) +
  guides(color = guide_legend(override.aes = list(linewidth =2),
                              nrow=2,
                              byrow=TRUE))
lagged_fig_a

### 2023-09-03 writing/manuscript.Rmd
update figs

ADD TEST POSITIVITY FIGURE
FIGURE OUT WHICH STATES HAVE LAGS

Another difference we see is in the months of the summer of 2021, where the increase in Covidestim estimates appears lagged in comparison to the probabilistic bias simulation intervals. The difference we see between the Covidestim estimates and probabilistic bias intervals is likely a result of the way these approaches treat incomplete testing. The focus of the probabilistic bias analysis is to correct for incomplete testing, and as such the method is sensitive to changes in the total number tested and the positivity rate. By contrast, while Covidestim models probabilities of diagnosis by symptom state to allow for variation in case ascertainment, the total number of tests is not an input into the model, so model estimates are not affected by changes in testing rate. This is particularly relevant during the summer of 2021, when an increase in test positivity preceded the increase in observed tests, leading to an increase in the probabilistic bias simulation intervals that precedes the increase in Covidestim estimates.

@@

@@

@@

>>>>

Another difference we see is in the months of the summer of 2021, where the increase in Covidestim estimates appears lagged in comparison to the probabilistic bias simulation intervals. The difference we see between the Covidestim estimates and probabilistic bias intervals is likely a result of the way these approaches treat incomplete testing. The focus of the probabilistic bias analysis is to correct for incomplete testing, and as such the method is sensitive to changes in the total number tested and the positivity rate (see Figure DIAGRAM in the methods section). By contrast, while Covidestim models probabilities of diagnosis by symptom state to allow for variation in case ascertainment, the total number of tests is not an input into the model, so model estimates are not affected by changes in testing rate. This is particularly relevant during the summer of 2021, when an increase in test positivity preceded the increase in observed tests, leading to an increase in the probabilistic bias simulation intervals that precedes the increase in Covidestim estimates in many states (Supplementary Figure lagged-figs). Differences attributed to how these methods consider test positivity are likely to be heightened during times where increases in infections outpace testing capacity or test utilization.

@@

# Discussion

@@

We acquired state-level PCR testing data from public-use data on HealthData.gov, which includes both positive and negative test counts [@unitedstatesdepartmentofhealthandhumanservices]. At the county level, we obtained publicly available data from Massachusetts published on the state government website [@departmentofpublichealth].

Covidestim estimates for dates after December 2021 were obtained from the Covidestim API, and those  for dates after December 2021 were obtained from the legacy page.

@@

# Data Availability

# Code Availability

### 2023-09-03 writing/supplement.Rmd
update figs

```{r lagged-figs, fig.cap="States where the Covidestim infection estimates are lagged relative to the probabilistic bias estimates, that is, they are most highly correlated when the Covidestim infections are lagged by one or more biweeks. (a) Simulation intervals demonstrating the lag seen around the delta peak in summer 2021. (b) Test positivity and positive tests for each state. Increases in test positivity tend to precede increases in positive tests, and these increases in test positivity result in increases in the estimates probduced by the probabilistic bias analysis seen in part (a)."}

>>>>

```{r lagged-figs, fig.cap="States where the Covidestim infection estimates appear to be lagged relative to the probabilistic bias estimates, that is, they are most highly correlated when the Covidestim infections are lagged by one or more biweeks. (a) Simulation intervals demonstrating the lag seen around the delta peak in summer 2021. (b) Test positivity and positive tests for each state. Increases in test positivity tend to precede increases in positive tests, and these increases in test positivity result in increases in the estimates probduced by the probabilistic bias analysis seen in part (a)."}
